Inequality in non-cognitive skills: a cross-cohort comparison

### 2018-09-03 codes/R/paper/paper_jpube.Rnw
current version of tables and figures

<<READ, cache=FALSE >>=
library(here)
read_chunk(here("codes/R/1_load.R"))
read_chunk(here("codes/R/2_efa.R"))
read_chunk(here("codes/R/3_inv_main.R"))
read_chunk(here("codes/R/3_inv_aux.R"))
read_chunk(here("codes/R/3_inv_fit.R"))
read_chunk(here("codes/R/4_scores.R"))
read_chunk(here("codes/R/4_scores_aux.R"))
read_chunk(here("codes/R/5_graphs.R"))
read_chunk(here("codes/R/5_graphs_ineq.R"))
read_chunk(here("codes/R/5_graphs_item.R"))
read_chunk(here("codes/R/6_selectvars.R"))
read_chunk(here("codes/R/8_regs_dep.R"))
read_chunk(here("codes/R/8_regs_outc.R"))
read_chunk(here("codes/R/99_auxfun.R"))

@@

<<FA_SCORES_BIND>>=
# bind scores with data
@

@@

<<SELECTVARS>>=
# define variables and formulas for regression and decomposition
@

<<REGS_DEP, dependson="SELECTVARS">>=
# run regressions explaining EXT INT
@

<<LOAD_DATA_OUTC, dependson="SELECTVARS">>=
# load outcomes data
@

<<REGS_OUTC, dependson="LOAD_DATA_OUTC">>=
# run regressions explaining adolescent/adult outcomes
@

@@

\begin{figure}
<<FACDENS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=10, fig.width=10, out.width='\\linewidth'>>=
@
\caption{Scores}
\justify {\scriptsize \emph{Notes}: Density of scores from final model.
\par}
\end{figure}

\begin{figure}
<<RAWHIST, dependson="FA_SCORES_BIND", dev='tikz', fig.height=10, fig.width=10, out.width='\\linewidth'>>=
@
\caption{Sum scores}
\justify {\scriptsize \emph{Notes}: Density of sum scores without factor analysis.
\par}
\end{figure}

@@

\begin{table}
\caption{Quantile differences in scores}\label{tab:iqdiff}
\centering\setstretch{1.2}\footnotesize
<<INEQ_STATS, dependson="FA_SCORES_BIND", results='asis'>>=

@@

\begin{figure}
<<FACINEQ_FSC, dependson="FA_SCORES_BIND", dev='tikz', fig.height=9, fig.width=10, out.width='\\linewidth'>>=
fi <- fi_fsc # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]],
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Inequality by father occupation -- normalised}
\end{figure}

\begin{figure}
<<FACINEQ_PS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=9, fig.width=10, out.width='\\linewidth'>>=
fi <- fi_ps # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]],
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Inequality by mother education -- normalised}
\end{figure}

\begin{figure}
<<FACINEQ_SM, dependson="FA_SCORES_BIND", dev='tikz', fig.height=9, fig.width=10, out.width='\\linewidth'>>=
fi <- fi_sm # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]],
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Inequality by mother smoking in pregnancy -- normalised}
\end{figure}

\begin{figure}
<<ITEMINEQ_PS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=8, fig.width=12, out.width='\\linewidth'>>=
itempl_ps
@
\caption{Item-level changes in inequality -- mother education}
\justify {\scriptsize \emph{Notes}: The graph displays the ratio between low and high SES in the prevalence of each behaviour, separately by cohort.
\par}
\end{figure}

\begin{figure}
<<ITEMINEQ_FSC, dependson="FA_SCORES_BIND", dev='tikz', fig.height=8, fig.width=12, out.width='\\linewidth'>>=
itempl_fsc
@
\caption{Item-level changes in inequality -- father occupation}
\justify {\scriptsize \emph{Notes}: The graph displays the ratio between low and high SES in the prevalence of each behaviour, separately by cohort.
\par}
\end{figure}

\begin{figure}
<<ITEMINEQ_SM, dependson="FA_SCORES_BIND", dev='tikz', fig.height=8, fig.width=12, out.width='\\linewidth'>>=
itempl_sm
@
\caption{Item-level changes in inequality -- mother smoking in pregnancy}
\justify {\scriptsize \emph{Notes}: The graph displays the ratio between low and high SES in the prevalence of each behaviour, separately by cohort.
\par}
\end{figure}

@@

# robust SEs
robust_se <- list()
robust_p  <- list()
for (i in 1:length(regs)) {
  cov1              <- sandwich::vcovHC(regs[[i]], type = "HC1")
  ct                <- lmtest::coeftest(regs[[i]], vcov = cov1)
  robust_se[[i]] <- ct[, 2]
  robust_p[[i]] <- ct[, 4]
}

texreg::texreg(regs,
               stars = c(0.1, 0.05, 0.01), digits = 3,
               override.se = robust_se, override.pvalues = robust_p,
               single.row = FALSE, scriptsize = TRUE,
               caption = "", label = "",
               omit.coef = omitcoefs,
               groups = grouplabs,
               custom.columns = list("p-value" = r_ext$Mp, "p-value" = r_ext$Fp), custom.col.pos = c(4,6),
               custom.coef.names = varlabs,
               custom.model.names = c("Males BCS", "Males MCS", "Females BCS", "Females MCS"),
               custom.note = "",
               include.rmse = FALSE,
               table = FALSE
               )

@@

# robust SEs
robust_se <- list()
robust_p  <- list()
for (i in 1:length(regs)) {
  cov1              <- sandwich::vcovHC(regs[[i]], type = "HC1")
  ct                <- lmtest::coeftest(regs[[i]], vcov = cov1)
  robust_se[[i]] <- ct[, 2]
  robust_p[[i]] <- ct[, 4]
}

texreg::texreg(regs,
               stars = c(0.1, 0.05, 0.01), digits = 3,
               override.se = robust_se, override.pvalues = robust_p,
               single.row = FALSE, scriptsize = TRUE,
               caption = "", label = "",
               omit.coef = omitcoefs,
               groups = grouplabs,
               custom.columns = list("p-value" = r_int$Mp, "p-value" = r_int$Fp), custom.col.pos = c(4,6),
               custom.coef.names = varlabs,
               custom.model.names = c("Males BCS", "Males MCS", "Females BCS", "Females MCS"),
               custom.note = "",
               include.rmse = FALSE,
               table = FALSE
               )

@@

\begin{table}
\centering\footnotesize
\caption{Predictors of adolescent and adult outcomes -- BCS}
<<REGS_OUTC_BCS, results='asis', dependson="REGS_OUTC">>=

# collapse to matrix
bcs_outctab_mat <- do.call(rbind,lapply(bcs_outctab[c("smktry16", "bmi16", "hnvq30", "lhgrpay38")],matrix,ncol=9,byrow=FALSE))
table <- bcs_outctab_mat
printoutctab(table, 4)

@@

\begin{table}
\centering\footnotesize
\caption{Predictors of adolescent outcomes -- MCS}
<<REGS_OUTC_MCS, results='asis', dependson="REGS_OUTC">>=

# collapse to matrix
mcs_outctab_mat <- do.call(rbind,lapply(mcs_outctab[c("smktry14", "bmi14")],matrix,ncol=9,byrow=FALSE))
table <- mcs_outctab_mat
printoutctab(table, 2)

@@

>>>>

<<READ, cache=FALSE >>=
library(here)
read_chunk(here("codes/R/1_load.R"))
read_chunk(here("codes/R/2_efa.R"))
read_chunk(here("codes/R/3_inv_main.R"))
read_chunk(here("codes/R/3_inv_aux.R"))
read_chunk(here("codes/R/3_inv_fit.R"))
read_chunk(here("codes/R/4_bootstrap.R"))
read_chunk(here("codes/R/5_selectvars.R"))
read_chunk(here("codes/R/6_graphs.R"))
read_chunk(here("codes/R/6_graphs_ineq.R"))
read_chunk(here("codes/R/6_graphs_item.R"))
read_chunk(here("codes/R/8_regs_dep.R"))
read_chunk(here("codes/R/8_regs_outc.R"))
read_chunk(here("codes/R/99_auxfun.R"))

@@

<<FA_BOOTSCORES_READ>>=
# read bootstrap data
@

<<SELECTDATA, dependson="FA_BOOTSCORES_READ">>=
# define data, variables and formulas for regression and decomposition
@

<<SELECTDATA_OUTC, dependson="SELECTDATA">>=
# load outcomes data
@

@@

<<REGS_DEP, dependson="SELECTDATA">>=
# run regressions explaining EXT INT
@

<<REGS_OUTC, dependson="SELECTDATA_OUTC">>=
# run regressions explaining adolescent/adult outcomes
@

@@

\begin{figure}
<<FACDENS, dependson="SELECTDATA", dev='tikz', fig.height=10, fig.width=10, out.width='\\linewidth'>>=
@
\caption{Scores}
\justify {\scriptsize \emph{Notes}: Density of scores from final model.
\par}
\end{figure}

\begin{figure}
<<RAWHIST, dependson="SELECTDATA", dev='tikz', fig.height=10, fig.width=10, out.width='\\linewidth'>>=
@
\caption{Sum scores}
\justify {\scriptsize \emph{Notes}: Density of sum scores without factor analysis.
\par}
\end{figure}

@@

\begin{table}
\caption{Quantile differences in scores}\label{tab:iqdiff}
\centering\setstretch{1.2}\footnotesize
<<INEQ_STATS, dependson="SELECTDATA", results='asis'>>=

@@

\begin{figure}
<<FACINEQ_FSC, dependson="SELECTDATA", dev='tikz', fig.height=9, fig.width=10, out.width='\\linewidth'>>=
fi <- fi_fsc # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]],
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Inequality by father occupation -- normalised}
\end{figure}

\begin{figure}
<<FACINEQ_PS, dependson="SELECTDATA", dev='tikz', fig.height=9, fig.width=10, out.width='\\linewidth'>>=
fi <- fi_ps # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]],
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Inequality by mother education -- normalised}
\end{figure}

\begin{figure}
<<FACINEQ_SM, dependson="SELECTDATA", dev='tikz', fig.height=9, fig.width=10, out.width='\\linewidth'>>=
fi <- fi_sm # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]],
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Inequality by mother smoking in pregnancy -- normalised}
\end{figure}

\begin{figure}
<<ITEMINEQ_FSC, dependson="SELECTDATA", dev='tikz', fig.height=9, fig.width=11, out.width='\\linewidth'>>=
itempl_fsc
@
\caption{Item-level changes in inequality -- father occupation}
\justify {\scriptsize \emph{Notes}: The graph displays the ratio between low and high SES in the prevalence of each behaviour, separately by cohort.
\par}
\end{figure}

\begin{figure}
<<ITEMINEQ_PS, dependson="SELECTDATA", dev='tikz', fig.height=9, fig.width=11, out.width='\\linewidth'>>=
itempl_ps
@
\caption{Item-level changes in inequality -- mother education}
\justify {\scriptsize \emph{Notes}: The graph displays the ratio between low and high SES in the prevalence of each behaviour, separately by cohort.
\par}
\end{figure}

\begin{figure}
<<ITEMINEQ_SM, dependson="SELECTDATA", dev='tikz', fig.height=9, fig.width=11, out.width='\\linewidth'>>=
itempl_sm
@
\caption{Item-level changes in inequality -- mother smoking in pregnancy}
\justify {\scriptsize \emph{Notes}: The graph displays the ratio between low and high SES in the prevalence of each behaviour, separately by cohort.
\par}
\end{figure}

@@

# robust SEs
robust_se <- list()
robust_p  <- list()
bootstrap_se  <- list(r_ext$BCS.M[["OLSbootse"]],r_ext$MCS.M[["OLSbootse"]],r_ext$BCS.F[["OLSbootse"]],r_ext$MCS.F[["OLSbootse"]])
for (i in 1:length(regs)) {
  cov1              <- sandwich::vcovHC(regs[[i]], type = "HC1")
  ct                <- lmtest::coeftest(regs[[i]], vcov = cov1)
  robust_se[[i]] <- ct[, 2]
  robust_p[[i]] <- ct[, 4]
}

texreg::texreg(regs,
               stars = c(0.1, 0.05, 0.01), digits = 3,
               override.se = bootstrap_se, override.pvalues = robust_p,
               single.row = FALSE, scriptsize = TRUE,
               caption = "", label = "",
               omit.coef = omitcoefs,
               groups = grouplabs,
               custom.columns = list("p-value" = r_ext$Mp, "p-value" = r_ext$Fp), custom.col.pos = c(4,6),
               custom.coef.names = varlabs,
               custom.model.names = c("Males BCS", "Males MCS", "Females BCS", "Females MCS"),
               custom.note = "",
               include.rmse = FALSE,
               table = FALSE
               )

@@

# robust SEs
robust_se <- list()
robust_p  <- list()
bootstrap_se  <- list(r_int$BCS.M[["OLSbootse"]],r_int$MCS.M[["OLSbootse"]],r_int$BCS.F[["OLSbootse"]],r_int$MCS.F[["OLSbootse"]])
for (i in 1:length(regs)) {
  cov1              <- sandwich::vcovHC(regs[[i]], type = "HC1")
  ct                <- lmtest::coeftest(regs[[i]], vcov = cov1)
  robust_se[[i]] <- ct[, 2]
  robust_p[[i]] <- ct[, 4]
}

texreg::texreg(regs,
               stars = c(0.1, 0.05, 0.01), digits = 3,
               override.se = bootstrap_se, override.pvalues = robust_p,
               single.row = FALSE, scriptsize = TRUE,
               caption = "", label = "",
               omit.coef = omitcoefs,
               groups = grouplabs,
               custom.columns = list("p-value" = r_int$Mp, "p-value" = r_int$Fp), custom.col.pos = c(4,6),
               custom.coef.names = varlabs,
               custom.model.names = c("Males BCS", "Males MCS", "Females BCS", "Females MCS"),
               custom.note = "",
               include.rmse = FALSE,
               table = FALSE
               )

@@

\begin{table}
\centering\footnotesize
\caption{Predictors of adolescent outcomes -- BCS}\label{tab:bcs_adol}
<<REGS_OUTC_BCS1, results='asis', dependson="REGS_OUTC">>=

# collapse to matrix
bcs_outctab_mat <- do.call(rbind,lapply(bcs_outctab[c("smktry16", "bmi16", "thn16", "owt16")],matrix,ncol=9,byrow=FALSE))
table <- bcs_outctab_mat
printoutctab(table, 4)

@@

\begin{table}
\centering\footnotesize
\caption{Predictors of adult outcomes -- BCS}\label{tab:bcs_adult}
<<REGS_OUTC_BCS2, results='asis', dependson="REGS_OUTC">>=

# collapse to matrix
bcs_outctab_mat <- do.call(rbind,lapply(bcs_outctab[c("hnvq30", "lhgrpay38", "obs42")],matrix,ncol=9,byrow=FALSE))
table <- bcs_outctab_mat
printoutctab(table, 3)

@@

\begin{table}
\centering\footnotesize
\caption{Predictors of adolescent outcomes -- MCS}\label{tab:mcs_adol}
<<REGS_OUTC_MCS, results='asis', dependson="REGS_OUTC">>=

# collapse to matrix
mcs_outctab_mat <- do.call(rbind,lapply(mcs_outctab[c("smktry14", "bmi14", "thn14", "owt14")],matrix,ncol=9,byrow=FALSE))
table <- mcs_outctab_mat
printoutctab(table, 4)

\clearpage
{\color{red} 
\textbf{Interpretation of relationship between early-life socioemotional skills and adolescent/adult outcomes} \\
Remember that a kid with good externalising skills is a kid that is NOT restless/hyperactive and does not have conduct problems (fighting, tantrums). (In fact in some factorisations of the Rutter, this factor separates into two separate factors similar to conduct and hyperactivity). A kid with good internalising skills is NOT neurotic, worried, fearful, or solitary.\\
In this light, starting from \autoref{tab:bcs_adol}: better EXT at 5 implies less smoking at 16 (a less `rebellious' child with less conduct problems?), and better INT at 5 implies more smoking at 16 (a more `social' child?). A similar thing is observed for BMI and overweight (although in the latter INT is not significant), but notice that this is limited to females -- more neurotic and restless females are larger. Almost none of the BCS children are thin/underweight at 16. In most of these adolescent outcomes, cognition does not make a difference.\\
A similar picture emerges for MCS in \autoref{tab:mcs_adol} -- note that here the age for the adolescent outcomes is 14 and not 16. But in MCS it seems that internalising skills matter less. There is some small indication that more neurotic females are more likely to be thin, but the prevalence is again really low. Also consider that all of these are LPM estimates.\\
For BCS adult outcomes (\autoref{tab:bcs_adult}), cognition dominates. But there is some significance for higher education in females: MORE neurotic (perhaps more focused on studies, care about outcomes) and LESS restless females have a higher likelihood of achieving a higher degree.
}
